---
title: "Team 6 Network Final Project: US Airports"
author: "Team 6: Jack Chu, Xijia Yuan, & Eva Wang"
date: "2023-05-03"
format:
  html:
    theme: united
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
setwd("~/Documents/Mod 4/Networks Theory and Analysis/Final Project/")
```


```{r, message=FALSE}
library(plyr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(sqldf)
library(openxlsx)
library(stringr)

library(ggraph)
library(ggplot2)
library(ggcorrplot)
library(mapview)
library(ggnetwork)

library(sna)
library(igraph) 
library(igraphdata)
library(network)

library(statnet)
library(btergm)
library(lme4)
library(lmerTest)
library(performance)
```

# Import Data

```{r}
data("USairports", package = "igraphdata")
nl <- get.data.frame(USairports, what = "vertices")
el <- get.data.frame(USairports, what = "edges")
el$logical <- which_loop(USairports)
el <- el[el$logical==FALSE,]
```

## Data Prep for ERGM

```{r}
new_edges <- el %>% 
  group_by(from, to) %>% 
  summarize(Sum_Distance=sum(Distance), Real_Distance=mean(Distance))


new_edges_1 <- el %>% 
  group_by(from) %>% 
  summarize(Sum_Passengers = sum(Passengers), Sum_Seats = sum(Seats), Sum_Depart=sum(Departures))

names(new_edges_1)[1] <- "name"

new_nodes <- merge(nl, new_edges_1, by="name", all.x = TRUE)

new_nodes[is.na(new_nodes)] <- 0

new_nodes$Sum_Passengers_scale <- scale(new_nodes$Sum_Passengers)[, 1]

new_nodes$Sum_Depart_scale <- scale(new_nodes$Sum_Depart)[, 1]

net_sna <- network(new_edges, matrix.type = "edgelist", 
                   directed = T, vertex.attr = new_nodes)

net_sna


un_net <- igraph::graph_from_data_frame(new_edges, directed = F, new_nodes)
new_edges$dup <- igraph::which_multiple(un_net)
new_edges2 <- new_edges %>% filter(dup == FALSE)
un_net_sna <- network(new_edges2, matrix.type = "edgelist",
                   directed = F, vertex.attr = new_nodes)
un_net_sna 


net <- graph_from_data_frame(el, directed = T, vertices = nl)
new_net <- graph_from_data_frame(new_edges, directed = T, vertices = new_nodes)
```

## Data Cleaning for MLM

```{r carrier region}
region <- read.xlsx("region.xlsx")

el2 <- merge(el, region, by = "Carrier")
el2 <- el2[,-9]

colnames(el2)[9:10] <- c("Carrier_Country", "Carrier_USstate")
```

```{r from & to}
city_split <- strsplit(nl$City, ",")
airport_code_location <- purrr::map_df(1:length(city_split), ~{
  data.frame(Code = nl$name[.x],
             Location = gsub(" ", "", city_split[[.x]][2]))
})

# Convert abb to names
state_code <- read.csv("state.csv")
airport_code_location$Name <- 
  ifelse(airport_code_location$Location %in% state_code$Code, state_code$State, 
         ifelse(airport_code_location$Location == "PR", "Puerto Rico",
                ifelse(airport_code_location$Location == "TT","Trinidad and Tobago",
                       ifelse(airport_code_location$Location == "VI", "U.S. Virgin Islands", NA))))
# airport_code_location[is.na(airport_code_location$Name),] # check

# merge w/el2
from_region <- merge(el2, airport_code_location, 
                     by.x = "from", by.y = "Code",)
el2$from_region <- from_region$Name

to_region <- merge(el2, airport_code_location, 
                     by.x = "to", by.y = "Code",)
el2$to_region <- to_region$Name

#write.csv(el2, "Final_el.csv", row.names = FALSE)
```

# Visualization

## Network Overall

```{r}
ggraph(net) + 
  geom_edge_link0(width = 0.05, alpha = .3) + 
  geom_node_point(size = 0.35) +
  ggnetwork::theme_blank()


airports <- read.csv('https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat', header = F)
# head(airports)

V(USairports)$lat=airports[match(V(USairports)$name, airports[,5]), 7] 
V(USairports)$long=airports[match(V(USairports)$name, airports[,5]), 8]

#remove loops and make undirected
usair <- as.undirected(simplify(USairports))

#remove airports whose codes didn't match the OpenFlights database (and hence returned "NA" for latitude)
usair <- delete.vertices(usair, which(is.na(V(usair)$lat)==TRUE))

#remove nodes in the Eastern and Southern Hemispheres (US territories). This will make the plot easier to see.
usair <- delete.vertices(usair, which(V(usair)$lat<0))
usair <- delete.vertices(usair, which(V(usair)$long>0))

#keep only the largest connected component of the network ("giant component"). this also makes the network easier to see.
decomp <- decompose.graph(usair)
usair <- decomp[[1]]

longlat=matrix(c(V(usair)$long, V(usair)$lat), ncol=2) #set up layout matrix 
par(mar=c(1,1,1,1))
plot(usair, layout=longlat, vertex.label="", vertex.size = 2)
```

## Visualization for locations

```{r}
# arrival airport
count <- table(el2$to_region)
to_region_viz <- data.frame(state = names(count),
                            count = as.numeric(count))

states <- read.csv("location.csv") # geocode lat and long for state and US territory
to_map_data <- merge(to_region_viz, states, 
                  by.x = "state", by.y = "STNAME")

pal = mapviewPalette("mapviewTopoColors")

mapview()+
  mapview(to_map_data, 
          xcol = "LONGITUDE", ycol = "LATITUDE", zcol = "count",
          cex = "count", 
          label = paste(to_map_data$state, ":", to_map_data$count),
          layer.name = c("# Arrival"),
          col.regions = pal(100), at = seq(min(to_map_data$count), max(to_map_data$count), 250),
          grid = FALSE, alpha = 1)
```


```{r}
# departure airport
countd <- table(el2$from_region)
from_region <- data.frame(state = names(countd),
                            count = as.numeric(countd))

states <- read.csv("location.csv") # geocode lat and long for state and US territory
from_map <- merge(from_region, states, 
                  by.x = "state", by.y = "STNAME")

mapview()+
  mapview(from_map, 
          xcol = "LONGITUDE", ycol = "LATITUDE", zcol = "count",
          cex = "count", 
          label = paste(from_map$state, ":", from_map$count),
          layer.name = c("# Depature"),
          col.regions = pal(100), at = seq(min(from_map$count), max(from_map$count), 250),
          grid = FALSE, alpha = 1)
```

```{r}
# Carrier location

counts <- table(el2$Carrier_USstate)
carrier_data <- data.frame(state = names(counts),
                          num = as.numeric(counts))
carrier_viz <- merge(carrier_data, states,
                     by.x = "state", by.y = "STNAME")

mapview()+
  mapview(carrier_viz, 
          xcol = "LONGITUDE", ycol = "LATITUDE", zcol = "num",
          cex = "num", 
          label = paste(carrier_viz$state, ":", carrier_viz$num),
          layer.name = c("# of carriers"),
          col.regions = pal(100), at = seq(min(carrier_viz$num), max(carrier_viz$num), 935),
          grid = FALSE, alpha = 1)
```

# Centrality

## Degree
```{r}
ggraph(net) + 
  geom_edge_link0(width = 0.05, color = "black", alpha = .3) +
  geom_node_point(fill = "green", color = "black", shape = 21, 
                  size = V(net)$size <- igraph::degree(net, mode = "all") / 300) +
  ggnetwork::theme_blank()

airport_deg <- as.data.frame(igraph::degree(net, mode = "all"))
names(airport_deg)[1] <- "degree"
airport_deg$airport <- row.names(airport_deg)
head(airport_deg[order(airport_deg$degree, decreasing = T),],5)
```

## Closeness
```{r}
comp_s <- igraph::components(net, mode = "strong")

max_comp <- which.max(comp_s$csize)

big_comp <- induced.subgraph(net, which(comp_s$membership == max_comp)) #subset the data to only include the largest connected component
big_comp

ggraph(net) + 
  geom_edge_link0(width = 0.05, color = "black", alpha = .3, 
                  arrow = grid::arrow(angle = 7.5,
                                      length = unit(0.08, "in"),
                                      type = "closed"), 
                  start_cap = circle(2, 'mm'),
                  end_cap = circle(2, 'mm')) +
  geom_node_point(fill = "gold", color = "black", shape = 21, 
                  size = igraph::closeness(net)*10) +
  ggnetwork::theme_blank()

ggraph(big_comp) + 
  geom_edge_link0(width = 0.05, color = "black", alpha = .3, 
                  arrow = grid::arrow(angle = 7.5,
                                      length = unit(0.08, "in"),
                                      type = "closed"), 
                  start_cap = circle(2, 'mm'),
                  end_cap = circle(2, 'mm')) +
  geom_node_point(fill = "gold", color = "black", shape = 21, 
                  size = igraph::closeness(big_comp)*5000) +
  ggnetwork::theme_blank()

airport_clo <- as.data.frame(igraph::closeness(big_comp))
names(airport_clo)[1] <- "closeness"
airport_clo$airport <- row.names(airport_clo)
head(airport_clo[order(airport_clo$closeness, decreasing = T),],10)
```

## Betweeness
```{r}
ggraph(net) + 
  geom_edge_link(width = 0.05, color = "black", alpha = .3, 
                  arrow = grid::arrow(angle = 10,
                                      length = unit(0.08, "in"),
                                      type = "closed"),
                  start_cap = circle(2, 'mm'),
                  end_cap = circle(2, 'mm')) +
  geom_node_point(fill = "pink", color = "black", shape = 21,
                  size = igraph::betweenness(net, directed = T)/17500) +
  ggnetwork::theme_blank()
airport_bet <- as.data.frame(igraph::betweenness(net, directed = T))
names(airport_bet)[1] <- "betweeness"
airport_bet$airport <- row.names(airport_bet)
head(airport_bet[order(airport_bet$betweeness, decreasing = T),],5)

# Edge Betweeness
# ggraph(net) + 
#   geom_edge_link(width = 0.05, color = "black", alpha = .3, 
#                   arrow = grid::arrow(angle = 10,
#                                       length = unit(0.08, "in"),
#                                       type = "closed"),
#                   start_cap = circle(2, 'mm'),
#                   end_cap = circle(2, 'mm'),
#                   aes(width = edge.betweenness(net, directed = T))) +
#   scale_edge_width(range = c(0.1, 1000)) +
#   geom_node_point(shape = 21, size = 0.5) + 
#   ggnetwork::theme_blank() + theme(legend.position = "none")
# 
# airport_ebet <- edge.betweenness(net, directed = T)
# names(airport_ebet)[1] <- "edge.betweeness"
```

# Structure

## Assortativity

Assortativity: is a measure of the tendency of nodes in a network to connect to other nodes that are similar or dissimilar to themselves in terms of certain characteristics or attributes.

```{r}
net_num <- graph_from_data_frame(new_edges, directed = T, vertices = new_nodes)

assortativity(net_num, V(net_num)$Sum_Passengers, directed = T)

assortativity(net_num, V(net_num)$Sum_Depart, directed = T)

assortativity_degree(net_num, directed = T)

# may be disassortative, but not significant, could be mixed
```

## Transitivity

Transitivity: ratio of the number of triangles that a node is part of to the maximum possible number of triangles it could be part of.

```{r}
trans <- igraph::transitivity(new_net, type = "local", isolates = "zero")
summary(trans)
```

## Efficiency

Efficiency: The average path between a node's neighbors, shorter path -> more efficient, high value will be this node can go to other place 

```{r}
local_efficiency(new_net)
average_local_efficiency(new_net)

new_nodes$local_transitivity <- trans 
new_nodes$local_efficientcy <- local_efficiency(new_net)

## Sort the dataframe
result1 <- sqldf("SELECT * FROM new_nodes ORDER BY local_efficientcy desc")
result2 <- sqldf("SELECT * FROM new_nodes ORDER BY local_transitivity desc")
```

## Diameter
```{r}
farthest_vertices(new_net)
```

## Distance
```{r}
mean_distance(new_net)
all_shortest_paths(new_net, from = "SBN", to  = "HNL")
```

## Triad
```{r}
sna::dyad.census(net_sna)
sna::triad.census(net_sna)
```

# ERGM

```{r}
model_master <- ergm(net_sna ~ edges + 
                 nodecov("Sum_Passengers_scale") +
                 absdiff("Sum_Passengers_scale") + nodecov("Sum_Depart_scale") + absdiff("Sum_Depart_scale") + mutual) 
summary(model_master)


gof.model1_btergm <- btergm::gof(model_master, nsim = 100, # checking goodness of fit of model estimates against 100 simulated networks
                                   statistics = c(esp, deg, odeg, ideg, geodesic, 
                                                  triad.directed)) #picked these statistics
plot(gof.model1_btergm)

plogis(coef(model_master)[['edges']] + coef(model_master)[['nodecov.Sum_Passengers_scale']] + 
         coef(model_master)[['absdiff.Sum_Passengers_scale']] + coef(model_master)[['nodecov.Sum_Depart_scale']] + coef(model_master)[['absdiff.Sum_Depart_scale']] + coef(model_master)[['mutual']])

mcmc.diagnostics(model_master)


Cug_Edges <- cug.test(net_sna, gtrans, cmode="edges")
plot(Cug_Edges)
```

## Visualization for ERGM
```{r}
ergm1 <- simulate(model_master)
ergm1_net <- ggraph(ergm1) +
  geom_edge_link0(color = "black", alpha = 0.15) +
  geom_node_point(fill = "#B0A498", color = "black", shape = 21) +
  labs(title = "Simulation 1") +
  ggnetwork::theme_blank() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
ergm1_net
```

# Multilevel Model

## Random intercept model

```{r}
# Correlation matrix
flight_corr <- el2[,-c(1:3, 10)]
flight_corr$Carrier_Country <- as.numeric(as.factor(flight_corr$Carrier_Country))
flight_corr$from_region <- as.numeric(as.factor(flight_corr$from_region))
flight_corr$to_region <- as.numeric(as.factor(flight_corr$to_region))
flight_corr_table <- cor(flight_corr)


ggcorrplot(flight_corr_table, method = c('square'), hc.order = TRUE, type = "lower", outline.color = 'white') +
  theme_minimal() + 
  labs(x = "", title = "Correlation Matrix", y = "") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        panel.background = element_blank(),
        axis.text.x = element_text(angle = 90),
        text = element_text(size = 15)) 
```

```{r}
# Remove highly correlated variable: Seats, Departures, to_region
mlm1 <- lmer(Passengers ~  Carrier_Country + 
               Aircraft + Distance + (1 | from_region), 
               data = el2, REML = T,
               na.action = na.omit)
summary(mlm1)

performance::icc(mlm1)
```

```{r}
el3 <- el2[el2$Carrier_Country=="USA",] # only focus on US carriers

mlm2 <- lmer(Passengers ~  Carrier_USstate + 
               Aircraft + Distance + (1 | from_region), 
               data = el3, REML = T,
               na.action = na.omit)
summary(mlm2)

performance::icc(mlm2)
```

